---
title: "Aides Drug"
author: "Jaci Heckman"
date: "Summer 2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
  
Our methods included obtaining all the categorical data, and finding out the number of drug users versus the information of infection rates. 

```{r}
druguser<- filter(triUmp, DrugUser=="yes")
```

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
```{r}
barchartGC(~infected + treatment,data=druguser)
barchartGC(~infected + treatment,data=druguser, type="percent")

```
    
The treatment information can be displayed neatly with two slides comparing the answer of whether or not a patient received the drug and their infection being altered. The first slide shows the frequency of the drug administration and the usage of the placebo versus the drug, the pink column shows that there is a greater frequency of patients given the placebo had the disease less affected. Whereas the blue column represents that the frequency was less when the drug was introduced to the patient with the disease. This drastic measurement is shown on slide two.  


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, data=druguser)
rowPerc(table1)
colPerc(table1)
```  

In this numerical summary the numbers of patients affected by the drug are placed blatantly to the patients affected by the placebo drug. The totals both remain at 100, but the comparative numbers of the success of the placebo patients are a 47.59, compared to the second portion under the drug patients are 97.11. These numbers can initially provide us that the results show a definite difference of numerical data.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The next data series shows three detailed tests, the results of which will be explained using one of them, the Ficher’s test. This includes the comparative number of infected patients and the likelihood of the results being the case. The probability was concluded in this test as a result of .00000000423, this number being marginally too small forces us to reject the null hypothesis and we are in favor of the inferential results. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
  
In this conclusion we were able to discover that the hypothesis was inherently correct. The drug when given to infected patients compared to ones given a placebo were showing graphic positive results. This was also displayed by showing two column slides that show apporximately 24% affected patients increased almost three times the amount by using the actual drug. This was verified by the numerical summary and the three segment test where we organized the Ficher's test to explain the provided data. 
